### 2013-02-14 doc/xps2013/proposal.tex
intro fine pass

\paragraph{Irregular applications frequently request small amounts of off-node data.} On multinode systems, the challenges presented by low locality are analogous, and exacerbated by the increased latency of going off-node. Irregular applications also present a challenge to mass market network technology, which is designed to transfer large blocks of data, not the word-sized references emitted by irregular application tasks. Injection rate into the network is insufficient to utilize wire bandwidth when blocks are below about two-kilobytes, so any straightforward communication strategy severely under-utilizes the network.

@@

\item \textbf{Reinvent network communication:} Our preliminary work has
revealed limitations of the current Grappa runtime system with regards to
network scaling. This proposal plans to enhance the network layer with new
forms of communication scheduling and aggregation, and explore new forms of
hardware/software interactions of remote direct memory access (RDMA).

@@

\item \textbf{Support for \ssd\ only memory systems:} The size of data being
processed by large systems today (hundreds of terabytes) dwarfs what can fit
into a single system. Multiple system nodes are required to maintain and
access all of this data efficiently. However, even with hundreds of channels
to persistent storage, loading this data into DRAM can prove a bottleneck. The
Grappa runtime is designed to make applications tolerate extremely high memory
latency, thus we will explore using solid state disks as main memory, 
further taking advantage of Grappa's latency tolerance properties.

@@

Efficiently executing irregular computations on commodity hardware is \emph{hard}.  Much about commodity hardware is seemingly designed for the opposite application characteristics: caches assume applications have locality; networks are built with low injection rates but high bandwidth, assuming large packet sizes; even runtimes and operating systems are built assuming relatively infrequent communication and synchronization compared to computation.  This proposal rests on the seemingly outlandish claim that this commodity hardware / software stack can be coerced, through another layer of software, into efficiently executing irregular applications that both lack locality and communicate frequently with small amounts of data.  A folksy way of saying it is we think we can jam a square peg through a round hole.  Fortunately we have some preliminary evidence we won't get wedged along the way.  This is an intellectually challenging endeavor, that we believe is well suited for the NSF XPS program.  In summary, our goal is to exploit parallelism from irregular computations to efficiently execute them at scale.

@@

Irregular computations are of increasing economic importance.  Moreover, they are key to several national security applications.  If commodity hardware can be cajoled into executing such applications efficiently it will open the door for new and rich applications.  Furthermore, irregular applications will then reap the benefits of advances in commodity processors and networks, rather than requiring custom hardware that advances only at much higher cost.  Graph queries, machine-learning and data-driven science will be enabled by the technology we aim to create from the ideas in this grant proposal.  In addition our work will include the construction of a comprehensive irregular benchmark suite, something to help guide our own and the field's research.

>>>>

\paragraph{Irregular applications frequently request small amounts of off-node
data.} On multinode systems (e.g., clusters of computers), the challenges
presented by low locality are analogous, and exacerbated by the increased
latency of going off-node. Irregular applications also present a challenge to
commodity network technology, which is designed to transfer large blocks of
data, not the word-sized references emitted by irregular application tasks.
Injection rate into the network is insufficient to utilize wire bandwidth when
blocks are below about two-kilobytes, so any straightforward communication
strategy severely under-utilizes the network.

@@

\item \textbf{Reinvent network communication:} Our preliminary work has
revealed limitations of the current Grappa runtime system with regards to
network scaling. This proposal plans to enhance the network layer with whole
new forms of communication scheduling and aggregation, and explore new forms
of hardware/software interactions of remote direct memory access (RDMA).

@@

\item \textbf{Support for \ssd\ only data structures:} The size of data being
processed by large systems today (hundreds of terabytes) dwarfs what can fit
into a single system. Multiple system nodes are required to maintain and
access all of this data efficiently. However, given the lack of locality in the application, loading this data into DRAM is wasteful, in addition to adding overheads. The
Grappa runtime is designed to make applications tolerate extremely high memory
latency, thus we will explore having program data structures reside in solid state disks only. This further takes advantage of Grappa's latency tolerance properties to deal with larger data sets and reduce system cost.

@@

Efficiently executing irregular computations on commodity hardware is a
\emph{difficult} problem. Much about commodity hardware is seemingly designed
for the opposite application characteristics: caches assume applications have
locality; networks are built with low injection rates but high bandwidth,
assuming large packet sizes; even runtimes and operating systems are built
assuming relatively infrequent communication and synchronization compared to
computation. This proposal rests on the seemingly outlandish claim that this
commodity hardware / software stack can be coerced, through another layer of
software, into efficiently executing irregular applications that both lack
locality and communicate frequently with small amounts of data. A folksy way
of saying it is we think we can jam a square peg through a round hole.
Fortunately we have some preliminary evidence we won't get wedged along the
way. This is an intellectually challenging endeavor, that we believe is well
suited for the NSF XPS program. In summary, our goal is to exploit parallelism
from irregular computations to efficiently execute them at scale.

@@

Irregular computations are of increasing economic importance.  Moreover, they are key to several national security applications.  If commodity hardware can be cajoled into executing such applications efficiently it will open the door for new and rich applications.  Furthermore, irregular applications will then reap the benefits of advances in commodity processors and networks, rather than requiring custom hardware that advances only at much higher cost.  Graph queries, machine-learning and data-driven science will significantly benefit from the technology we aim to create from the ideas in this grant proposal.  In addition our work will include the construction of a comprehensive irregular benchmark suite, something to help guide our own and the field's research.
